---
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
    includes:
      in_header: preamble.tex
fontsize: 11pt
geometry: margin=1in
linestretch: 1
---

\begin{flushleft}
\Large \textbf{STSCI 4100: IMDb Analysis Q5} \\
\normalsize Tony Oh (do256), William Rhee (wr86) \\
\normalsize May 16, 2025
\end{flushleft}

```{r setup, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
library(data.table)
library(mgcv)
gc()
full_dt <- readRDS("full_dt.rds")
```

## QUESTION 5:
An extra question we want to ask is: 
**does experience in creative roles improve film ratings?**

In previous questions, we explored whether a film having a specified **writer** 
(Question 2) impacts ratings or whether the **number of votes** (Question 1) 
impacts ratings. But now we want to ask a deeper question: 
**Does the experience of the writers, directors, or actors improve a film’s average rating?**

We defined experience as the **number of films** an individual has previously 
been in with a specific role. For each film, we compute the 
**median experience** of each group.

We start by reshaping `principals` for an easier count of a person's 
experience. Then, we compute the **median experience** per film, and merge it 
into our dataset.

```{r q5-1, message=FALSE, echo = FALSE}
# [1] Compute experience metrics for writers, directors, and actors

q5_dt <- full_dt
long_principals <- full_dt[, .(tconst, nconst, category)]
long_principals <- long_principals[
  , .(nconst = unlist(nconst), 
      category = unlist(category)), 
  by = tconst
]

# Count experience per person per category
role_experience <- long_principals[, .N, by = .(nconst, category)]
setnames(role_experience, "N", "experience")

# Computes median experience per film for role
get_role_experience_by_film <- function(role_name) {
  role_exp <- role_experience[category %in% role_name]
  
  role_map <- full_dt[, .(tconst, nconst, category)]
  role_map <- role_map[, .(
    nconst = unlist(nconst),
    category = unlist(category)
  ), by = tconst]
  role_map <- role_map[category %in% role_name]
  
  role_map <- merge(role_map, role_exp, by = "nconst", all.x = TRUE)
  role_exp_by_film <- role_map[, .(
    exp_median = median(experience, na.rm = TRUE)
  ), by = tconst]
  
  role_name <- if ("actor" %in% role_name) "actor" else role_name[1]
  setnames(role_exp_by_film, "exp_median", paste0(role_name, "_exp_median"))
  return(role_exp_by_film)
}

# Get experience for each group
writer_exp_by_film <- get_role_experience_by_film(c("writer"))
director_exp_by_film <- get_role_experience_by_film(c("director"))
actor_exp_by_film <- get_role_experience_by_film(c("actor", "actress"))

# Merge into q5_dt
q5_dt <- merge(q5_dt, writer_exp_by_film,   by = "tconst", all.x = TRUE)
q5_dt <- merge(q5_dt, director_exp_by_film, by = "tconst", all.x = TRUE)
q5_dt <- merge(q5_dt, actor_exp_by_film,    by = "tconst", all.x = TRUE)

# Replace NAs with 0
exp_cols <- grep("_exp_median$", names(q5_dt), value = TRUE)
for (col in exp_cols) {
  q5_dt[is.na(get(col)), (col) := 0]
}
```

We fit two GAMs:  
- One casual with separate effects for each group  
- One with an interaction between writer and director experience

```{r q5-2}
# [2] Fit GAMs
# Median
gam_median <- gam(averageRating ~ 
                    s(writer_exp_median) +
                    s(director_exp_median) +
                    s(actor_exp_median),
                  data = q5_dt)

# Median, Interaction
gam_median_inter <- gam(averageRating ~ 
                          te(writer_exp_median, director_exp_median),
                        data = q5_dt)
```

```{r q5-3, echo = FALSE}
# [3] Visualize the effects of experience
par(mfrow = c(1, 3))
plot(gam_median, select = 1, main = "Writer Experience (Median)")
plot(gam_median, select = 2, main = "Director Experience (Median)")
plot(gam_median, select = 3, main = "Actor Experience (Median)")
```

Each of the three plots shows how the median experience in one group affects 
the rating prediction while holding the other groups constant.

```{r q5-4, echo = FALSE}
# 3D interaction plot (writer, director)
par(mfrow = c(1, 1))
vis.gam(gam_median_inter,
        view = c("writer_exp_median", "director_exp_median"),
        plot.type = "persp",
        color = "topo",
        theta = 30, phi = 30,
        main = "Writer, Director Median Experience")
```

This plot shows the interaction between writer median experience and director 
median experience and how it affects the ratings prediction. 

Here is what we can takeaway from our analysis: 

  + **Writer experience** shows a positive trend. 
  + **Director experience** shows a nonlinear shape, where more director 
  experience tends to decrease the ratings predictions. 
  + **Actor experience** has a mostly flat effect. 
  + The **interaction plot** reveals that a moderate level of combined 
  experience gives good ratings predictions rather than high levels of both. 

These trends are statistically significant due to large sample size, but 
practically unreasonable to consider convincing. Still, we seem to be able to 
see that writer experience may be important in a film's success. 
